Inference for multivariate peaks over threshold models

### 2021-10-12 conclusion.tex
update

In this paper, we have built upon the definition of the multivariate Pareto described in \cite{ferreira2014}
  to establish a useful parametric representation of its dependence structure, through describing
  the distribution of its angular component; it being supported on the positive orthant of the unit
  hypersphere under the $\mathcal{L}_{\infty}$ norm, $\mathcal{S}_{\infty}^{d-1}$.  We then demonstrated an
  inherent difficulty in establishing a distribution supported in this space.  To build that representation,
  we presented a distribution based on a product of independent Gammas, projected onto a supporting manifold
  $\mathcal{S}_{p}^{d-1}$, which will converge to $\mathcal{S}_{\infty}^{d-1}$ as $p\to\infty$.  We then
  established a means of model comparison on $\mathcal{S}_{\infty}^{d-1}$ through the use of a negative
  definite kernel defined in that space.

>>>>

\section{Conclusion\label{sec:conclusion}}
In this paper, we have built upon the definition of the multivariate Pareto described in \cite{ferreira2014}
  to establish a useful parametric representation of its dependence structure, through describing
  the distribution of its angular component; it being supported on the positive orthant of the unit
  hypersphere under the $\mathcal{L}_{\infty}$ norm, ${\mathbb S}_{\infty}^{d-1}$.  We then demonstrated an
  inherent difficulty in establishing a distribution supported in this space.  To build that representation,
  we presented a distribution based on a product of independent Gammas, projected onto a supporting manifold
  ${\mathbb S}_{p}^{d-1}$, which will converge to ${\mathbb S}_{\infty}^{d-1}$ as $p\to\infty$.  We then
  established a means of model comparison on ${\mathbb S}_{\infty}^{d-1}$ through the use of a negative
  definite kernel defined in that space.

### 2021-10-12 evaluation.tex
update

\section{Model Comparison on the Hypersphere}
\label{sec:evaluation}
It is not immediately obvious what criteria to use to judge these models and decide which best
  represents the data's generating distribution.  As the task resides on $\mathcal{S}_{\infty}^{d-1}$,
  our options are somewhat limited in terms of what model criteria we can use.  Because of the aforementioned
  difficulty in evaluating a density directly in this space, we choose to use model evaluation
  criteria that can operate on samples from the posterior predictive distribution.  For this reason,
  we have opted to use the \emph{posterior predictive loss} criterion of \cite{gelfand1998} and the
  \emph{energy score} criterion of \cite{gneiting2007}.  Both of these metrics require calculating
  some distance in the target space, and this section will be devoted to that end.  As both these
  metrics operate on the posterior predictive distribution for each observation, we will also be
  including a metric operating on the overall posterior predictive distribution, in the form of a
  modified Kullbeck-Liebler divergence.

\subsection{Posterior Predictive Loss Criterion}
The posterior predictive loss criterion, \emph{PPL} is introduced in \cite{gelfand1998}.  When we
  assume a squared error loss function, then for the $i$th observation, the posterior predictive
  loss criterion is computed as
  \begin{equation}
    \label{eq:ppl}
    D_{k}^{(i)} = \text{Var}\left(X_i\right) + \frac{k}{k + 1}\left(\text{E}[X_i] - x_i\right)^2,
  \end{equation}
  where $X_i$ is a random variable from the posterior predictive distribution for $x_i$.  The
  scalar $k$ is a weighting factor by which we arbitrarily scale the importance of goodness of fit
  relative to precision.  In our analysis, we take the limit as $k\to\infty$, weighting both
  parts equally.  Interpreting this criterion, a smaller $\text{Var}(\bm{X}_{\ell})$ indicates a higher
  precision, and a smaller $(\text{E}[\bm{X}_{\ell}] - \bm{x}_{\ell})^2$  indicates a better fit.  Thus,
  smaller is better.  Note that this is defined for a univariate distribution.  As we are dealing
  with a multivariate distribution, we will be taking the posterior predictive loss
  summed over all dimensions.  That is,
  \begin{equation}
    \label{eq:ppl2}
    D_k^{(i)} = \frac{1}{d}\sum_{\ell = 1}^{d}\left[\text{Var}(X_{i,\ell}) +
                  \frac{k}{k+1}\left(\text{E} [X_{i,\ell}] - x_{i,\ell}\right)^2\right]
  \end{equation}
  Then we report the average $D_k$, taken over all $i$.

\subsection{Energy Score}
The energy score of Gneiting, et al\cite{gneiting2007}, is a generalization of the continuous ranked
  probability score, or \emph{crps}, defined for a multi-dimensional random variable.
  \begin{equation}
    \label{eq:es}
    \text{ES}\left(P, \bm{x}_i\right) =  \text{E}_p g\left(\bm{X}_i, \bm{x}_i\right) -
                                    \frac{1}{2}\text{E}_p g\left(\bm{X}_i,\bm{X}_i^{\prime}\right)
  \end{equation}
  where $g(\cdot)$ identifies a negative definite kernel, and $\bm{X}_i^{\prime}$ represents another
  replicate from the posterior predictive distribution of $\bm{x}_i$. A function $g$ is a negative definite kernel if it is symmetric in its arguments, $g(x_1,x_2) = g(x_2,x_1)$, and for which
  $\sum_{i =1}^n\sum_{j=1}^n\alpha_i\alpha_jg(\bm{x}_1,\bm{x}_2) \leq 0$ for all positive integers n, with the
  restriction that $\sum_{i=1}^n\alpha_i = 0$.  Euclidean distance is one example of a negative definite kernel,
  and the one most often used.  However, on $\mathcal{S}_p^{d-1}$ for $p > 1$, Euclidean distance
  can under-report the actual distance required for travel between two points. That distortion is
  maximized on $\mathcal{S}_{\infty}^{d-1}$.

\subsection{Distance on the positive orthant of the Hypercube}
  \label{subsec:distance}
The positive orthant of the unit hypercube, defined in Euclidean geometry, is that structure for
  which, in a given point on the hypercube, all dimensions of that point are between 0 and 1, and
  at least one dimension must be 1.  Developing terminology, we can consider observations for which
  the $j$th dimension is equal to 1, to be on the $j$th \emph{face}.  The intersection of the $i$th
  and $j$th face is a $d-2$ dimensional cube, and observations in this space have dimensions $i,j$
  equal to 1.

A \emph{distance} in this space is a \emph{geodesic} on this space. From geometry, we know that
  the geodesic, or shortest path between 2 points along the surface of a $d$ dimensional figure
  corresponds to at least one \emph{unfolding}, or \emph{rotation} of the $d$-dimensional
  figure into a $d-1$ dimensional space.  The appropriate term for the structure generated by this
  unfolding is a \emph{net}.  For the appropriate net, a line segment connecting the two points and
  staying within the boundaries of the net corresponds to the shortest path between those points\citep{pappas1989}, and is thus a geodesic.  The length of that line segment is properly
  defines the distance required for travel between those points.

Consider a 3-dimensional cube.  Consider 2 points on this 3-dimensional cube,
  $\bm{a}_1 = (x_1,y_1,z_1)$, and $\bm{a}_2 = (x_2,y_2,z_2)$. Let's say that the two points are
  on the same face.  Then the distance between those two points, the distance one has to travel
  along the space to move from one point to the other, is calculated by Euclidean norm.  Now,
  consider two points on separate faces.  All faces are pairwise adjacent, as we have stated, so
  in order to move to the other point, we must \emph{at least} move to the intersection between the
  faces, then to the other point.  Let $\bm{a}_1$ lie on the $x$ face, and $\bm{a}_2$ lie on the
  $y$ face.  That is, $\bm{a}_1 = (1, y_1, z_1)$, $\bm{a}_2 = (x_2, 1, z_2)$.  Then traveling
  between these points we must at least pass through the intersection of faces $x,y$.  One possible
  net representation of this is unfolding the $y$ face alongside the $x$ face.  We accomplish this
  by applying a rotation and translation to $a_2$, corresponding to the following:
  \begin{equation}
    \label{eq:1drotation}
    a_2^{\prime} = \begin{bmatrix}
    1 \\
    2 \\
    0
    \end{bmatrix}
    +
    \begin{bmatrix}
    0  & 0 & 0 \\
    -1 & 0 & 0 \\
    0  & 0 & 1
    \end{bmatrix}
    \begin{bmatrix}
    x_2 \\
    1 \\
    z_2 \\
    \end{bmatrix} = \begin{bmatrix}
    1 \\
    2 - x_2 \\
    z_2
    \end{bmatrix}
  \end{equation}
  Then, if this is the appropriate net, the distance from $\bm{a}$ to $\bm{b}$ becomes
  $\lVert \bm{a}_1 - \bm{a}_2^{\prime}\rVert_2$. However, there
  is another possible net we must consider, travelling first through the $z$ face then to the
  $y$ face.  The rotation for that becomes:
  \begin{equation}
    \label{eq:2drotation}
    a_2^{\prime} = \begin{bmatrix}
    1 \\
    2 \\
    2
    \end{bmatrix}
    +
    \begin{bmatrix}
    0 & 0 & 0  \\
    0 & 0 & -1 \\
    -1 & 0 & 0
    \end{bmatrix}
    \begin{bmatrix}
    x_2 \\
    1 \\
    z_2
    \end{bmatrix} = \begin{bmatrix}
    1 \\
    2 - z_2 \\
    2 - x_2
    \end{bmatrix}
  \end{equation}
Every successive rotation is relative to the last face.  So, as the number of dimensions grows, the
  number of possible rotations grows as well.  As we have $d$ faces, if 2 observations are on
  different faces, then there are $\sum_{j = 1}^{d-2}\binom{d-2}{j} + 1$ possible rotations to
  consider\footnote{There are truly $d!$ possible nets, but when we consider starting and ending
  faces fixed, and that portions of the net that diverge after the ending face are irrelevant,
  we arrive at that number of rotations that we need consider}.  While this is not
  insurmountable, it is computationally costly. However, for a valid energy score, a true distance
  is not necessary.  What is needed is a negative definite kernel.

\begin{prop}
For points $a,b \in \mathcal{S}_{\infty}^{d-1}$ a valid negative definite kernel can be formed as
  \begin{equation}
    g(\bm{a},\bm{b}) = \begin{cases}
        \pnorm{\bm{b}-\bm{a}}{2} &\text{ if }\argmax_{\ell}\bm{a} = \argmax_{\ell}\bm{b}\\
        \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} &\text{ otherwise}
    \end{cases}
  \end{equation}
  where $\bm{c}$ resides on the intersection between the faces of $\bm{a}$ and $\bm{b}$, and minimizes $g(\bm{a},\bm{b})$.
\end{prop}

In the first case, where $\bm{a}$ and $\bm{b}$ reside on the same face, then that is trivially a
  negative definite kernel.  In the second case, where $\bm{a}$ and $\bm{b}$ reside on separate
  faces, we can be assured of symmetry in the arguments as $\bm{c}$ is unique, and $\pnorm{\bm{c} - \bm{a}}{2} = \pnorm{\bm{a} - \bm{c}}{2}$.  Then for the other property of negative definiteness,
  \begin{equation*}
    \begin{aligned}
      \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j g(\bm{x}_1,\bm{x}_2) &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j \bigg[\pnorm{\bm{c} - \bm{x}_1}{2} + \pnorm{\bm{x}_2 - \bm{c}}{2}\bigg]\\
      &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{c} - \bm{x}_1}{2} + \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{x}_2 - \bm{c}}{2}
    \end{aligned}
  \end{equation*}
  which is less than $0$ in both terms by definition.

To compute this energy score as described would require optimization of the interim point $\bm{c}$.
  If we again consider the one-dimensional rotation in Equation~\ref{eq:1drotation}---we
  can view that as a numerically easier analogue accomplishing the same goal: the distance from
  the starting point, to some optimal point along the intersection between the starting and
  ending faces, to the ending point.  The strait line will pass through the optimal point, so
  the optimization of the interim point is accomplished \emph{en passant}.

>>>>

\section{Scores for distributions in ${\mathbb S}_\infty^{d-1}$\label{sec:evaluation}}

As explained in~\ref{subsubsec:norm}, we can not establish a \emph{density} on ${\mathbb S}_{\infty}^{d-1}$, restricting our ability to construct model selection criteria to sample-based approaches.  In \cite{gneiting2007}, Theorem~4 states that a score,
    \begin{equation}
    \label{eq:es}
    \text{S}\left(P, \bm{x}_i\right) =  \text{E}_p g\left(\bm{X}_i, \bm{x}_i\right) -
                                    \frac{1}{2}\text{E}_p g\left(\bm{X}_i,\bm{X}_i^{\prime}\right),
    \end{equation}
where $g(\cdot)$ is a negative definite kernel, is a \emph{proper} scoring rule.  A real valued function $g$ is a negative definite kernel if it is symmetric in its arguments, and $\sum_{i=1}^n\sum_{j=1}^na_ia_jg(x_i,x_j)\leq 0$ for all positive integers $n$, $a_1,\ldots,a_n\in{\mathbb R}$, and $\sum_{i = 1}^na_i = 0$. \cite{gneiting2007} introduces this quantity as the \emph{energy score}.

In Euclidean space, the conditions for a negative definite kernel are satisfied by Euclidean distance.  On ${\mathbb S}_{\infty}^{d-1}$, however, Euclidean distance will tend to under-estimate actual distance required for travel between two points.  Developing terminology, consider ${\mathbb C}_{\ell}^{d-2} = \lbrace x : x_{\ell} = 1\rbrace$ the \emph{$\ell$th face}.  Then for points on the same face, Euclidean distance will report the length of the shortest path.  For points on different faces, Euclidean distance will under-report that length.

The space, ${\mathbb S}_{\infty}^{d-1}$ exists as the limit of a sequence of manifolds, ${\mathbb S}_p^{d-1}$ for $p\to\infty$.  For a given manifold, the length of the shortest path between two points is a \emph{geodesic}; as the length of the geodesic is symmetric, and by definition satisfies the triangle inequality, we call it a \emph{distance} in that manifold.  Ideally, we would want to establish such a distance in ${\mathbb S}_{\infty}^{d-1}$, but as the surface is not differentiable, routines for establishing geodesics are not available.  However, as ${\mathbb S}_{\infty}^{d-1}$ is a portion of of a $d$-cube, we can borrow a result from geometry \citep{pappas1989} in that the length of the shortest path between 2 points on a geometric figure corresponds to the length of a straight line drawn between the points on an appropriate unfolding, rotation, or \emph{net} of the figure from $d$-dimensional to $d-1$-dimensional space.  The optimal net will have the shortest straight line between the points, so long as that line does not exit the net. As ${\mathbb S}_{\infty}^{d-1}$ has $d$ faces---each face pairwise adjacent, there are $d!$ possible nets.  However, as we are only interested in nets that begin and end on the source and destination faces respectively, and are not concerned with divergence after the destination face.  That reduces the number of nets for consideration to $\sum_{k = 0}^{d-2}\binom{d-2}{k}$.  This is still computationally burdensome for a large number of dimensions.

Instead, we develop as our kernel an upper bound on distance between points on separate faces as the length of the path from the source point, to an optimal point on the intersection between faces, to the destination point.
\begin{prop}
For points $a,b \in {\mathbb S}_{\infty}^{d-1}$ a valid negative definite kernel can be formed as
  \begin{equation}
    g(\bm{a},\bm{b}) = \begin{cases}
        \pnorm{\bm{b}-\bm{a}}{2} &\text{ if }\argmax_{\ell}\bm{a} = \argmax_{\ell}\bm{b}\\
        \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} &\text{ otherwise}
    \end{cases}
  \end{equation}
  where $\bm{c}$ resides on the intersection between the faces of $\bm{a}$ and $\bm{b}$, and minimizes $g(\bm{a},\bm{b})$.
\end{prop}

{\em Proof:}
In the first case, where $\bm{a}$ and $\bm{b}$ reside on the same face, then that is trivially a
  negative definite kernel.  In the second case, where $\bm{a}$ and $\bm{b}$ reside on separate
  faces, we can be assured of symmetry in the arguments as $\bm{c}$ is unique, and $\pnorm{\bm{c} - \bm{a}}{2} = \pnorm{\bm{a} - \bm{c}}{2}$.  Then for the other property of negative definiteness,
  \begin{equation*}
    \begin{aligned}
      \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j g(\bm{x}_1,\bm{x}_2) &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j \bigg[\pnorm{\bm{c} - \bm{x}_1}{2} + \pnorm{\bm{x}_2 - \bm{c}}{2}\bigg]\\
      &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{c} - \bm{x}_1}{2} + \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{x}_2 - \bm{c}}{2},
    \end{aligned}
  \end{equation*}
  which is less than $0$ in both terms.$\hfill\Box$

Evaluating $g$ as described requires optimization of the interim point $\bm{c}$; a $d-2$-dimensional optimization problem.  However, let us consider a computationally easier analogue: the aforementioned \emph{net}, where we move directly from the source face to the target face in one jump. Let $\bm{a} \in {\mathbb C}_{\ell}^{d-2}$, and $\bm{b} \in {\mathbb C}_{\jmath}^{d-2}$.  The rotation required to move the $\jmath$th face along the $\ell$th axis is
\begin{equation}
    \bm{b}^{\prime} = 
    \begin{cases}
        b_{i} &\text{for }i\neq \jmath,\ell\\
        1 &\text{for }i = \ell\\
        2 - b_{\ell} &\text{for }i = \jmath
    \end{cases}
\end{equation}
where $\cdot^{\prime}$ indicates a rotated point.  Then $g(\bm{a},\bm{b}) = \pnorm{\bm{a} - \bm{b}^{\prime}}{2}$.  By fitting the straight line between $\bm{a}$ and $\bm{b}^{\prime}$, the optimization of $\bm{c}$ is accomplished \emph{en passant}.

% {\bf After a sentence or two motivating the need for scoring distributions in ${\mathbb S}_\infty^{d-1}$ that can be calculated with a sample-based approach, mention the paper of \cite{gneiting2007} on proper scoring rules, and introduce the energy score. Cite the result in that paper that gives conditions for the energy score to be a proper scoring rule. Then discuss the problem of finding a metric in ${\mathbb S}_\infty^{d-1}$ that corresponds to a negative definite kernel. That is, the material in Section 4.3. Introduce and discuss the idea, the write the proposition, with its proof. I don't know if the example in the 3-dimensional cube is needed or not. Finally, say how the score will be calculated in our case.
% }
% It is not immediately obvious what criteria to use to judge these models and decide which best
%   represents the data's generating distribution.  As the task resides on ${\mathbb S}_{\infty}^{d-1}$,
%   our options are somewhat limited in terms of what model criteria we can use.  Because of the aforementioned
%   difficulty in evaluating a density directly in this space, we choose to use model evaluation
%   criteria that can operate on samples from the posterior predictive distribution.  For this reason,
%   we have opted to use the \emph{posterior predictive loss} criterion of \cite{gelfand1998} and the
%   \emph{energy score} criterion of \cite{gneiting2007}.  Both of these metrics require calculating
%   some distance in the target space, and this section will be devoted to that end.  As both these
%   metrics operate on the posterior predictive distribution for each observation, we will also be
%   including a metric operating on the overall posterior predictive distribution, in the form of a
%   modified Kullbeck-Liebler divergence.

% \subsection{Posterior Predictive Loss Criterion}
% The posterior predictive loss criterion, \emph{PPL} is introduced in \cite{gelfand1998}.  When we
%   assume a squared error loss function, then for the $i$th observation, the posterior predictive
%   loss criterion is computed as
%   \begin{equation}
%     \label{eq:ppl}
%     D_{k}^{(i)} = \text{Var}\left(X_i\right) + \frac{k}{k + 1}\left(\text{E}[X_i] - x_i\right)^2,
%   \end{equation}
%   where $X_i$ is a random variable from the posterior predictive distribution for $x_i$.  The
%   scalar $k$ is a weighting factor by which we arbitrarily scale the importance of goodness of fit
%   relative to precision.  In our analysis, we take the limit as $k\to\infty$, weighting both
%   parts equally.  Interpreting this criterion, a smaller $\text{Var}(\bm{X}_{\ell})$ indicates a higher
%   precision, and a smaller $(\text{E}[\bm{X}_{\ell}] - \bm{x}_{\ell})^2$  indicates a better fit.  Thus,
%   smaller is better.  Note that this is defined for a univariate distribution.  As we are dealing
%   with a multivariate distribution, we will be taking the posterior predictive loss
%   summed over all dimensions.  That is,
%   \begin{equation}
%     \label{eq:ppl2}
%     D_k^{(i)} = \frac{1}{d}\sum_{\ell = 1}^{d}\left[\text{Var}(X_{i,\ell}) +
%                   \frac{k}{k+1}\left(\text{E} [X_{i,\ell}] - x_{i,\ell}\right)^2\right]
%   \end{equation}
%   Then we report the average $D_k$, taken over all $i$.

% \subsection{Energy Score}
% The energy score of Gneiting, et al\cite{gneiting2007}, is a generalization of the continuous ranked
%   probability score, or \emph{crps}, defined for a multi-dimensional random variable.
%   \begin{equation}
%     \text{ES}\left(P, \bm{x}_i\right) =  \text{E}_p g\left(\bm{X}_i, \bm{x}_i\right) -
%                                     \frac{1}{2}\text{E}_p g\left(\bm{X}_i,\bm{X}_i^{\prime}\right)
%   \end{equation}
%   where $g(\cdot)$ identifies a negative definite kernel, and $\bm{X}_i^{\prime}$ represents another
%   replicate from the posterior predictive distribution of $\bm{x}_i$. A function $g$ is a negative definite kernel if it is symmetric in its arguments, $g(x_1,x_2) = g(x_2,x_1)$, and for which
%   $\sum_{i =1}^n\sum_{j=1}^n\alpha_i\alpha_jg(\bm{x}_1,\bm{x}_2) \leq 0$ for all positive integers n, with the
%   restriction that $\sum_{i=1}^n\alpha_i = 0$.  Euclidean distance is one example of a negative definite kernel,
%   and the one most often used.  However, on ${\mathbb S}_p^{d-1}$ for $p > 1$, Euclidean distance
%   can under-report the actual distance required for travel between two points. That distortion is
%   maximized on ${\mathbb S}_{\infty}^{d-1}$.

% \subsection{Distance on the positive orthant of the Hypercube}
%   \label{subsec:distance}
% The positive orthant of the unit hypercube, defined in Euclidean geometry, is that structure for
%   which, in a given point on the hypercube, all dimensions of that point are between 0 and 1, and
%   at least one dimension must be 1.  Developing terminology, we can consider observations for which
%   the $j$th dimension is equal to 1, to be on the $j$th \emph{face}.  The intersection of the $i$th
%   and $j$th face is a $d-2$ dimensional cube, and observations in this space have dimensions $i,j$
%   equal to 1.

% A \emph{distance} in this space is a \emph{geodesic} on this space. From geometry, we know that
%   the geodesic, or shortest path between 2 points along the surface of a $d$ dimensional figure
%   corresponds to at least one \emph{unfolding}, or \emph{rotation} of the $d$-dimensional
%   figure into a $d-1$ dimensional space.  The appropriate term for the structure generated by this
%   unfolding is a \emph{net}.  For the appropriate net, a line segment connecting the two points and
%   staying within the boundaries of the net corresponds to the shortest path between those points\citep{pappas1989}, and is thus a geodesic.  The length of that line segment is properly
%   defines the distance required for travel between those points.

% Consider a 3-dimensional cube.  Consider 2 points on this 3-dimensional cube,
%   $\bm{a}_1 = (x_1,y_1,z_1)$, and $\bm{a}_2 = (x_2,y_2,z_2)$. Let's say that the two points are
%   on the same face.  Then the distance between those two points, the distance one has to travel
%   along the space to move from one point to the other, is calculated by Euclidean norm.  Now,
%   consider two points on separate faces.  All faces are pairwise adjacent, as we have stated, so
%   in order to move to the other point, we must \emph{at least} move to the intersection between the
%   faces, then to the other point.  Let $\bm{a}_1$ lie on the $x$ face, and $\bm{a}_2$ lie on the
%   $y$ face.  That is, $\bm{a}_1 = (1, y_1, z_1)$, $\bm{a}_2 = (x_2, 1, z_2)$.  Then traveling
%   between these points we must at least pass through the intersection of faces $x,y$.  One possible
%   net representation of this is unfolding the $y$ face alongside the $x$ face.  We accomplish this
%   by applying a rotation and translation to $a_2$, corresponding to the following:
%   \begin{equation}
%     \label{eq:1drotation}
%     a_2^{\prime} = \begin{bmatrix}
%     1 \\
%     2 \\
%     0
%     \end{bmatrix}
%     +
%     \begin{bmatrix}
%     0  & 0 & 0 \\
%     -1 & 0 & 0 \\
%     0  & 0 & 1
%     \end{bmatrix}
%     \begin{bmatrix}
%     x_2 \\
%     1 \\
%     z_2 \\
%     \end{bmatrix} = \begin{bmatrix}
%     1 \\
%     2 - x_2 \\
%     z_2
%     \end{bmatrix}
%   \end{equation}
%   Then, if this is the appropriate net, the distance from $\bm{a}$ to $\bm{b}$ becomes
%   $\lVert \bm{a}_1 - \bm{a}_2^{\prime}\rVert_2$. However, there
%   is another possible net we must consider, travelling first through the $z$ face then to the
%   $y$ face.  The rotation for that becomes:
%   \begin{equation}
%     \label{eq:2drotation}
%     a_2^{\prime} = \begin{bmatrix}
%     1 \\
%     2 \\
%     2
%     \end{bmatrix}
%     +
%     \begin{bmatrix}
%     0 & 0 & 0  \\
%     0 & 0 & -1 \\
%     -1 & 0 & 0
%     \end{bmatrix}
%     \begin{bmatrix}
%     x_2 \\
%     1 \\
%     z_2
%     \end{bmatrix} = \begin{bmatrix}
%     1 \\
%     2 - z_2 \\
%     2 - x_2
%     \end{bmatrix}
%   \end{equation}
% Every successive rotation is relative to the last face.  So, as the number of dimensions grows, the
%   number of possible rotations grows as well.  As we have $d$ faces, if 2 observations are on
%   different faces, then there are $\sum_{j = 1}^{d-2}\binom{d-2}{j} + 1$ possible rotations to
%   consider\footnote{There are truly $d!$ possible nets, but when we consider starting and ending
%   faces fixed, and that portions of the net that diverge after the ending face are irrelevant,
%   we arrive at that number of rotations that we need consider}.  While this is not
%   insurmountable, it is computationally costly. However, for a valid energy score, a true distance
%   is not necessary.  What is needed is a negative definite kernel.

% \begin{prop}
% For points $a,b \in {\mathbb S}_{\infty}^{d-1}$ a valid negative definite kernel can be formed as
%   \begin{equation}
%     g(\bm{a},\bm{b}) = \begin{cases}
%         \pnorm{\bm{b}-\bm{a}}{2} &\text{ if }\argmax_{\ell}\bm{a} = \argmax_{\ell}\bm{b}\\
%         \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} &\text{ otherwise}
%     \end{cases}
%   \end{equation}
%   where $\bm{c}$ resides on the intersection between the faces of $\bm{a}$ and $\bm{b}$, and minimizes $g(\bm{a},\bm{b})$.
% \end{prop}

% In the first case, where $\bm{a}$ and $\bm{b}$ reside on the same face, then that is trivially a
%   negative definite kernel.  In the second case, where $\bm{a}$ and $\bm{b}$ reside on separate
%   faces, we can be assured of symmetry in the arguments as $\bm{c}$ is unique, and $\pnorm{\bm{c} - \bm{a}}{2} = \pnorm{\bm{a} - \bm{c}}{2}$.  Then for the other property of negative definiteness,
%   \begin{equation*}
%     \begin{aligned}
%       \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j g(\bm{x}_1,\bm{x}_2) &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j \bigg[\pnorm{\bm{c} - \bm{x}_1}{2} + \pnorm{\bm{x}_2 - \bm{c}}{2}\bigg]\\
%       &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{c} - \bm{x}_1}{2} + \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{x}_2 - \bm{c}}{2}
%     \end{aligned}
%   \end{equation*}
%   which is less than $0$ in both terms by definition.

% To compute this energy score as described would require optimization of the interim point $\bm{c}$.
%   If we again consider the one-dimensional rotation in Equation~\ref{eq:1drotation}---we
%   can view that as a numerically easier analogue accomplishing the same goal: the distance from
%   the starting point, to some optimal point along the intersection between the starting and
%   ending faces, to the ending point.  The strait line will pass through the optimal point, so
%   the optimization of the interim point is accomplished \emph{en passant}.
